---
title: "TJ Feature Engineering"
author: "Tim Morales"
date: "10/24/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/")
df<-read.csv("Fully Joined Df.csv")
```

Im going to drop a lot of the tommy john info as its either irrelevant or repeated elsewhere. 

Dropping and Flagging Variables 
```{r}
## removing random surgery information 
colnames(df[, c(185:228)])
df<-df[, -c(185:228)]

# make yearID and TJ a factor
df$yearID <- as.factor(df$yearID)
df$TJ<-as.factor(df$TJ)
levels(df$TJ) = c("No", "Yes")


#make a made postseason flag 
df$made_postseason = as.factor(ifelse(is.na(df$post_IPouts),0,1))

#i can fill playoff NAs with 0 
#colnames(df[,c(162:186)])
df[,c(162:186)][is.na(df[c(162:186)])] <- 0

#Making a pitchs per game stat 
df$pitches_per_game<- (df$pitch_count / df$p_game)

#Percentages if NA set to 0 
formatted.names<- c("n_ff_formatted","n_sl_formatted","n_ch_formatted","n_cukc_formatted","n_sift_formatted","n_fc_formatted","n_kn_formatted","n_fs_formatted", "n_fastball_formatted","n_breaking_formatted","n_offspeed_formatted")
df[formatted.names][is.na(df[formatted.names])] <- 0


```


Previous paper shows warm high schools increase risk. We use birth place and assign based on states in paper as well as handselected warm countries. 

```{r}
#warm weather indicator 
warmstates<- c("CA","AZ","NM","TX","LA","MS","AL","GA","FL","SC")
outsideUSCold<-c("Lithuania","Germany","CAN", "USA")
#print all countries 
levels(as.factor(df$birthCountry))

df$warm_birth_place <- ifelse(df$birthState %in% warmstates,1,0)
#make "notin"
`%notin%` <- Negate(`%in%`)
#outside US 
df$warm_birth_place <- as.factor(ifelse(df$birthCountry %notin% outsideUSCold,1,df$warm_birth_place))
```


```{r}
library(tidyverse)
library(lubridate)
#drop the 1 missing player nameFirst
### check miss to be fixed 
# colSums(is.na(df))
#i drop the below columns because theyre either not usefully or impossible to impute
df<- df %>% 
   mutate(debut = ymd(debut)) %>% 
  mutate_at(vars(debut), funs(year, month, day))%>%
  select(-c("meatball_swing_percent", "deathYear","deathMonth","deathDay","deathCountry","deathState","deathCity", "nameLast.y","nameFirst.y","X","X.1","birthMonth","birthDay","birthCountry","birthState","birthCity","nameGiven", "finalGame", "day", "debut", "n"))

#make the debut variable into year and month of debut 
df <- rename(df, c("year_debut" = "year" , "month_debut"= "month"))
df1<-df
```

Bin variables with lots of missing 
```{r}

df.not.bin<- df

col.w.na = colSums(is.na(df)) > 0
 

library(OneR)

col.w.na.names<-colnames(df)[col.w.na]
#col.w.na.names

#i make the function i want 
bin.omit.false<- function(x){
  bin(x, nbins = 4, na.omit = F, labels = c("Bottom_25", "Q2", "Q3", "Top_25")) }

doesnt.throw<- function(x){
  x<-ifelse(is.na(x)==T, "Does_Not_Throw", x)
}

df[col.w.na.names]<- lapply(df[col.w.na.names], bin.omit.false)

```

Feature engineer making these per inning stats 

```{r}
df$pitch_per_out <- (df$pitch_count/ df$IPouts)
df$innings_per_game <- (df$IPouts/ df$G.x)
df$pitcher_per_BF <- df$pitch_count/ df$BFP
df$batters_per_inning <- df$BFP / (df$IPouts/3)


df$fastball_per_inning <- df$pitch_count_fastball / (df$IPouts/3)
df$breakingball_per_inning <- df$pitch_count_breaking / (df$IPouts/3)
```

We remove nearzero variane variables unless they are related to previous research like IP or BFP

```{r}
library(caret)
nzv <- nearZeroVar(df)

colnames(df)[nzv]

nzv.final <- nzv[-c(11,14,15,25,33)]

df<- df[,-nzv.final]

```



One hot encode the factors using recipe 

```{r}

# Drop identifying character variables 
df.onehot <- df %>%
  select(- c("nameLast.x", "nameFirst.x", "PlayerName", "playerID", "teamID", "retroID", "bbrefID"))




library(recipes)
rec = recipe( ~ ., data = df.onehot)
rec_2 = rec %>% 
  step_dummy(all_predictors(),-all_numeric(), one_hot = T)
d_prep=rec_2 %>% prep(training = df.onehot, retain = T)

df.juice<-juice(d_prep)



write.csv(df.juice, "/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/ModelReadyData.csv")

```













