---
title: "Tommy John Project Data Cleaning"
author: "Tim Morales"
date: "9/24/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning

```{r}
library(tidyverse)
```

## Tommy John List 


```{r}
tj_list<-read.csv('/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/TJ List.csv')
table(tj_list$Level)
tj_list$Year_Mon<-as.numeric(paste(tj_list$Year,tj_list$Month,sep=""))


#keep everyone up to aug 2015 
mlb_only_tj_list<-subset(tj_list, Level=="MLB" & Year_Mon >= 20158)
mlb_only_tj_list<-subset(mlb_only_tj_list, Level=="MLB" & Year >= 2015)
table(mlb_only_tj_list$Level)


#create a variable to denote the last year before their surgery 
#if they got surgery after or during aril  in that year
# we can use the same season. 
#if before april, we use year prior.

mlb_only_tj_list$Year_for_stats<-ifelse(mlb_only_tj_list$Mon >= 3, mlb_only_tj_list$Year,mlb_only_tj_list$Year - 1) 


#if surgery done in 2020, auto use 2019
mlb_only_tj_list$Year_for_stats<-ifelse(mlb_only_tj_list$Year==2020, 2019, mlb_only_tj_list$Year_for_stats)
```


Then we must look at players who repeated surgery. If they have undergone surgery twice, put that in a new column, with binary 1 0, 1 deonting repeated. We also feature the column for those with repeated surgery how many years between surgery.

```{r}
df_repeated<- read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/Repeat TJ List.csv")
df_repeated<-df_repeated[!is.na(df_repeated$Years.Between),]
#merge together 
mlb_only_tj_list<-left_join(mlb_only_tj_list,df_repeated[,c(1,4)], by = "Player")

mlb_only_tj_list$Repeat_Injury<-ifelse(is.na(mlb_only_tj_list$Years.Between),0,1)

#then lets keep only the pitchers 
mlb_only_tj_list<-mlb_only_tj_list%>%
  filter(Position == "P")

```

Merge the names

I have a player list from Tanner Bell, MLB author. 
```{r}
translator.list<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/SFBB Player ID Map - PLAYERIDMAP.csv")
translator.list$fgid<-translator.list$IDFANGRAPHS
translator.list$bbrefid<-translator.list$IDPLAYER
```

I can now use left_join to add the bbrefIDs to the TJ list using fangraphs ID. 


There are still a few that do not have bbrefID, for them I will look up on my own and store in csv, then import. 

```{r}
tj.list.Id<-left_join(mlb_only_tj_list,translator.list[,c(43,44)], by = "fgid")

missing.indx<-which(is.na(tj.list.Id$bbrefid))
missing.names<-tj.list.Id$Player[missing.indx]
missing.names

hand.filled.id<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/manual fill bbref for missing.csv")[,c(1,2)]

#fill those missing values that I hand filled 
tj.list.Id$bbrefid[missing.indx]<-hand.filled.id$bbrefid

#just checking 
#tj.list.Id[,c(1,47)]
```

So now for me to be able to merge all of these labels together I will do so by baseball reference ID as it appears in all data sets. 

```{r}
#LAHMAN stats 
pitching <- read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/baseballdatabank-master/core/Pitching.csv")

#only keep after 2015 and making summing stats for each player each year
pitching<-pitching%>%
  filter(yearID >= 2015)
   
#get team with most innings played for when played for multiple teams
player.teams<-pitching[,c(1,2,4,5,13)]%>%
    group_by(playerID, yearID)%>%
   arrange(by_group=desc(IPouts))%>%
   filter(row_number() == 1)

pitching<-pitching[,-c(3,4,5,19,20)]
pitching<-aggregate(.~playerID+yearID,pitching, sum)

#add the character variables back on DOESNT WORK I NEED TO ONLY GET FIRST
pitching<-left_join(pitching, player.teams[,-5], by = c('playerID', 'yearID'))

#import names translation 
namelist<- read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/baseballdatabank-master/core/People.csv")

#inner join to get each player its needed labels 
named_pitching<- inner_join(pitching, namelist, by= 'playerID')

```

post season pitching

I dont have to be concerned with two teams in same year here but i need to combine each series for each player
```{r}
post.season.pitching<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/baseballdatabank-master/core/PitchingPost.csv")

#dropping calculated stats and characters like league and playoff round
post.season.pitching<-post.season.pitching[,-c(3,4,5,19,20)]

#combine stats across series 
post.season.pitching<-post.season.pitching%>%
  filter(yearID >= 2015)%>%
   group_by(playerID, yearID)%>%
   summarise_each(funs(sum))
               
               
```

The next step is to left join, adding injury information to players but only on the years being used as stats. 

Now we need to work with Sabermetrics datasets. 
MINIMUM 10 PA IN THESE LOWEST REQUIREMENT

```{r}
SM2015<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/2015SM.csv")
SM2016<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/2016SM.csv")
SM2017<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/2017SM.csv")
SM2018<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/2018SM.csv")
SM2019<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/2019SM.csv")


#combine into one df 
SMALL<-rbind(SM2015,SM2016,SM2017,SM2018,SM2019)
colnames(SMALL)[1]<-"nameLast"
colnames(SMALL)[2]<-"nameFirst"
colnames(SMALL)[3]<-"yearID"
```

Try and correctly merge 
```{r}
#make full name 
SMALL$PlayerName<- paste(SMALL$nameFirst,SMALL$nameLast,sep=" ")

#remove leading and trailing
SMALL$PlayerName<-trimws(SMALL$PlayerName, which = c("both"), whitespace = "[ \t\r\n]")

#NAMED PITCHING CLEANING 

named_pitching$nameFirst<-gsub("\\. ","\\.",named_pitching$nameFirst)

named_pitching$PlayerName<-paste(named_pitching$nameFirst,named_pitching$nameLast,sep=" ")

#remove leading and trailing

named_pitching$PlayerName<-trimws(named_pitching$PlayerName, which = c("both"), whitespace = "[ \t\r\n]")

```



```{r}
#first go to see who is missing
ij.pitch<-inner_join(SMALL,named_pitching, by = c("PlayerName","yearID"))

#we will track who is missing and manually fill w bbref id 
overlap<-SMALL$PlayerName %in% ij.pitch$PlayerName 
namecon<-data.frame(PlayerName=SMALL$PlayerName[overlap==F], yearID=SMALL$yearID[overlap==F])

# this just shows that these names are NOT the same between the two datasets.
# namecon$name %in% SMALL$PlayerName
# namecon$name %in% named_pitching$PlayerName
```

```{r}
#import overlap baseball reference numbers I hand filled 
SM.fill<-read.csv("/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/nameissues.csv")

namecon$bbrefID<-SM.fill$bbrefID

SM.fill.test<-left_join(namecon, named_pitching, by=c("bbrefID", "yearID"))
colnames(SM.fill.test)[1] = "PlayerName"

misfits<-anti_join(SMALL,named_pitching, by = c("PlayerName","yearID"))

rj.pitch<-right_join(SM.fill.test,misfits, by = c("PlayerName", "yearID"))
rj.pitch<-rj.pitch[,-52]
df.final<-rbind(ij.pitch,rj.pitch)

# colnames(rjpitch) %in% colnames(ij.pitch)

#making sure information is proper 
#removing duplicates 

df.final<-unique(df.final)

#looking for missing 
missing.inform<-is.na(df.final$p_game)
# 
# df.final$PlayerName[missing.inform==T]

#inconsistency in counting stat to confirm correct 
inconsist<-df.final$p_game == df.final$G
table(inconsist)

### we have to drop these people due to inconsistency in data 
df.final$PlayerName[inconsist==F]

## drop those 
df.pitch<-df.final[inconsist,]
```

In all, I have 3645 observations out of 3651 eligble season retained. Those dropped had either no Lahman data or could not be confirmed as reliable information. 

I now need to add the postseason information which I can do by player and playerID and yearID

```{r}
colnames(post.season.pitching)[-c(1,2)]<-paste("post",colnames(post.season.pitching)[-c(1,2)], sep ="_" )

post.season.pitching$bbrefID = post.season.pitching$playerID

df.plus.post<-left_join(df.pitch,post.season.pitching, by = c("playerID", "yearID"))

```

Finally, I have to add the tommy john information where needed 

```{r}
colnames(tj.list.Id)[47]<-"bbrefID"
colnames(tj.list.Id)[44]<-"yearID"

#make my TJ flag
tj.list.Id$TJ<-1
tj.list.Id[,-c(22:42)]

colnames(df.plus.post)[160]<-"bbrefID"

#able to keep 97 players with TJ 
df.total<-left_join(df.plus.post, tj.list.Id, by = c( "bbrefID", "yearID"))

#create the flag 
df.total$TJ<-ifelse(is.na(df.total$TJ), 0, 1)

```


We will require that pitchers face at least 30 batters in that season, going off of michigan papers' 10 innings. We lose 5 TJ players. and over 200 regular players. 

```{r}

injured <- subset(df.total, df.total$TJ == 1)
which(injured$BFP<30)
injured$PlayerName[which(injured$BFP<30)]

#keep only those with more than 30 batters faced 
df.total <- subset(df.total, df.total$BFP >= 30)

#write.csv(df.total, "/Users/timmorales/Desktop/STAT 6341/Tommy John Project/Data/Fully Joined Df.csv")
```




